---
title: "Live Session 6 Case Study Practice"
author: "Matthew Thacker and Damon Resnick"
date: "October 19, 2016"
output:
  html_document:
    keep_md: yes
  pdf_document: default
  word_document: default
---

This is a live session assignment and practice for the Case Study Assignemnt I.

The goal is to learn more about reproducable research by getting and cleaning data, creating an RStudio project,
and doing analysis with documentation that will help the researcher and other researchers in the future.

This project will include a README file, a data directory, an analysis directory, a paper directory, and a link to the an online repository in GitHub.

The data was taken from the Rolling Housing Sales for NYC:

http://www1.nyc.gov/home/search/index.page?search-terms=Rolling+sales+update

Our assignment was to look at the data for the borough of Manhattan.

The original code that this project is based off of was presented by the

#### Original Author: Benjamin Reddy

Taken from pages 49-50 of O'Neil and Schutt

require(gdata)

require(plyr) #Added by Monnie McGee

Install the gdata and plyr packages and set working directory in R.

```{r, message=FALSE}
library(plyr)
library(gdata)
setwd("C:/Users/hp/Desktop/SMU/Doing Data Science/Live Classes/Unit 7")
```
.

You need a perl interpreter to load an xls file in Windows.
It's automatic in Mac
mn <- read.xls("rollingsales_brooklyn.xls",pattern="BOROUGH")


So we will save the file as a csv and use read.csv instead:

```{r}
mn <- read.csv("./data/rollingsales_manhattan.csv",skip=4,header=TRUE)
```
.


#### Check the data:

```{r}
head(mn, 3)
tail(mn, 3)
summary(mn)
str(mn) # This is a very handy function in R! Displays the structure of an object.
```
.


Next up is to clean/format the data with regular expressions.  More on these later.
For now, know that the pattern "[^[:digit:]]" refers to members of the variable name that start with digits.
We will use the gsub command to replace them with a blank space.

We then create a new variable that is a "clean" version of sale.price; we make sale.price.n numeric, not a factor.

#### This means that we replaced the -'s with NA and made a new column called SALE.PRICE.N so that R can deal with the data better.
The gsub() function is useful for finding and replacing characters.
Counting the data use the count() function helps to tell how many numeric values and NAs there are in SALE.PRICE.N.

```{r}
mn$SALE.PRICE.N <- as.numeric(gsub("[^[:digit:]]","", mn$SALE.PRICE))
count(is.na(mn$SALE.PRICE.N))
```
.


The names() function makes all variable names lower case.

```{r}
names(mn) <- tolower(names(mn))
```
.


#### Get rid of leading digits
The frst 2 lines of the code below will again replace '-' with NA in R so that missing data will not throw off the analysis.

```{r}
mn$gross.sqft <- as.numeric(gsub("[^[:digit:]]","", mn$gross.square.feet))
mn$land.sqft <- as.numeric(gsub("[^[:digit:]]","", mn$land.square.feet))
mn$year.built <- as.numeric(as.character(mn$year.built))
```
.


#### Here we do a bit of exploration to make sure there's not anything weird going on with sale prices
Attaching 'mn' means that the commands after will automatically reference this data without needing to specify 'mn' for each command.

```{r}
attach(mn)
hist(sale.price.n) #Something is wrong! The histogram shows that most sales seem to be close to $0.
detach(mn)
```
.


#### Keep only the actual sales
[columns,rows]
This will create a data table called mn.sale from mn.  This identifies 0's (no given price) within the sale.price column and converts the entire observation to NA. This allows R to ignore those observations when performing analysis on the sale data.

```{r}
mn.sale <- mn[mn$sale.price.n!=0,]
```
.


#### Create a scatterplot of the mn.sale data when compared to sq. footage.

```{r}
plot(mn.sale$gross.sqft,mn.sale$sale.price.n)
```
.


#### Perform a log base 10 transformation of the data and create a new scatterplot.

```{r}
plot(log10(mn.sale$gross.sqft),log10(mn.sale$sale.price.n))
```
.


#### Let's look at 1, 2, and 3-family homes
This creates a new data table called mn.homes, which provides a list of family homes.
It accomplishes this by searching for the term 'family' within mn.sale building class and passes it to mn.homes.

```{r}
mn.homes <- mn.sale[which(grepl("FAMILY",mn.sale$building.class.category)),]
```
.


Look at the dimensions for the mn.homes object (observations and variables)

```{r}
dim(mn.homes)
```
.


#### Plot the log10 of sale.price vs. gross.sqft and look at the summary of sale.price

```{r}
plot(log10(mn.homes$gross.sqft),log10(mn.homes$sale.price.n))
summary(mn.homes[which(mn.homes$sale.price.n<100000),])
```
.


#### Remove outliers that seem like they weren't actual sales and show the plot

```{r}
mn.homes$outliers <- (log10(mn.homes$sale.price.n) <=5) + 0
mn.homes <- mn.homes[which(mn.homes$outliers==0),]
plot(log(mn.homes$gross.sqft),log(mn.homes$sale.price.n))
```
.


#### Create a .csv file called "cleaned_rollingsales_manhattan.csv" to save the cleaned data

```{r}
write.csv(mn.homes, "cleaned_rollingsales_manhattan.csv", row.names = TRUE)
```



